```{r, include = FALSE, cache = FALSE}
library(knitr)
# opts_knit$set(out.format = "latex")
knit_theme$set(knit_theme$get("greyscale0"))

options(replace.assign=FALSE,width=50)

opts_chunk$set(fig.path='figure/graphics-', 
               cache.path='cache/graphics-', 
               fig.align='center', 
               dev='pdf', fig.width=5, fig.height=5, 
               fig.show='hold', cache=FALSE, par=TRUE)
knit_hooks$set(crop=hook_pdfcrop)
suppressPackageStartupMessages(library(dplyr))
```

## Simple Hypothesis Testing

  Method A
  ------- ------- ------- ------- ------- -------
  78.64   79.01   79.57   79.52   80.71   79.95
  78.50   79.10   81.98   80.09   80.29   80.22
  ------- ------- ------- ------- ------- -------

  Method B
  ------- ------- ------- ------- ------- -------
  81.92   81.12   82.47   82.86   82.89   82.45
  82.51   81.11   83.07   82.77   82.38   83.14
  ------- ------- ------- ------- ------- -------

1.  We conducted an experiment and collected the data in the tables above. This
 data set isn't paired.^[I intentionally didn't make the data
    available for download so you would have to think about how to enter the
    data. You could enter it either Excel and import or directly into R.]
    a) Input the data into ^[Here I would suggest input the data into Excel and using `read_csv()`]. Combine the two data sets into a single data frame.

    ```{r}
    ##Data for question 1
    ## Easier using Excel and export as CSV
    x = c(78.64,79.01, 79.57, 79.52, 80.71, 79.95, 78.50,
      79.10, 81.98, 80.09, 80.29, 80.22)
    y = c(81.92, 81.12, 82.47, 82.86, 82.89, 82.45,
         82.51, 81.11, 83.07, 82.77, 82.38, 83.14)
    dd = data.frame(x, y)
    
    ```
    
    ```{r, echo = FALSE}
    d1 = data.frame(value = x)
    d2 = data.frame(value = y)
    ```
    
    ```{r, eval=FALSE}
    ## Suppose you have two separate data files. Here is some code that will help ## you combine them. First we read in the separate files:
    d1 = read.csv("Method1.csv")
    d2 = read.csv("Method2.csv")
    ```
    
    ```{r}
    ## In order to combine the data frames,
    ## they must have the same column names:
    head(d1, 2)
    head(d2, 2)
    ```
    
    ```{r}
    ## We combine data frames using rbind (row bind)
    d = rbind(d1, d2)
    ```
    
    ```{r}
    ## Finally we create a new column to indicate the Method
    ## rep is the replicate function. See ?rep
    d$Method = rep(1:2, each=12)
    head(d, 2)
    ```

    b) Exploratory data analysis. Construct boxplots, histograms and q-q plots for both data sets. Work
    out the means and standard deviations. Before carrying out any statistical
    test, what do you think your conclusions will be? Do you think the variances
    are roughly equal? Do you think the data conforms to a normal
    distribution.
    c) Carry out a two sample $t$-test. Assume that the variances are unequal.

    ```{r}
    t.test(value ~ Method, data=d, var.equal=FALSE)
    ```

    How does this answer compare with your intuition?

    d) Carry out a two sample $t$-test, assuming equal variances.

    ```{r}
    t.test(value ~ Method, data=d, var.equal=TRUE)
    ```

2. Suppose we are interested whether successful business
  executives are affected by their zodiac sign. We have collected 4265 samples
  and obtained the following data
  
    ```{r, echo = TRUE, results = "show"}
    data(zodiac, package = "jrAnalytics")
    head(zodiac)
    ```
    
    a) Carry out a $\chi^2$ goodness of fit test on the zodiac data. Are business executives distributed uniformly across zodiac signs?

    ```{r}
    x = zodiac$count
    m = chisq.test(x)
    ##Since p > 0.05 we can't accept the alternative hypothesis.
    ##However, the question is worded as though we can "prove" the Null
    ##hypotheis, which we obviously can't do.
    ```


    b) What are the expected values for each zodiac sign?

    ```{r}
    ##expected values
    (expected = m[["expected"]])
    ```


    c) The formula for calculating the residuals ^[These residuals are
    called Pearson residuals. Hint: use `str(m)` to extract the residuals.] is given by

    $$
    \frac{\text{observed} - \text{expected}}{\sqrt{\text{expected}}}
    $$

    Which residuals are large?

    ```{r}
    ##Residuals
    m[["residuals"]]
    ```

## One way ANOVA tables

1. A pilot study was developed to investigate whether music
influenced exam scores. Three groups of students listened to 10 minutes of
Mozart, silence or heavy metal before an IQ test. The results of the IQ test
are as follows

  ----------- --- --- --- --- --- --- ---
  Mozart      109 114 108 123 115 108 114
  Silence     113 114 113 108 119 112 110
  Heavy Metal 103  94 114 107 107 113 107
  ----------- --- --- --- --- --- --- ---
  
  a) Construct a one-way ANOVA table. Are there differences between treatment groups?

    ```{r}
    x1 = c(109, 114, 108, 123, 115, 108, 114)
    x2 = c(113, 114, 113, 108, 119, 112, 110)
    x3 = c(103, 94, 114, 107, 107, 113, 107)
    dd = data.frame(values = c(x1, x2, x3), type = rep(c("M", "S", "H"), each=7))
    m = aov(values ~ type, dd)
    summary(m)
    ##The p value is around 0.056.
    ##This suggests a difference may exist.
    ```

   b) Check the standardised residuals of your model.

    ```{r fig.keep="none"}
    plot(fitted.values(m), rstandard(m))
    ## Residual plot looks OK
    
    ```

   c) Perform a multiple comparison test to determine where the difference lies.
    
    ```{r}
    TukeyHSD(m)
    ```
